On the formation of H$\alpha$ line emission around classical T Tauri stars

### 2006-01-24 TTau_Halpha/kurosawa.tex
cut out bipolar stuff - th
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@630 25464f21-6099-402c-b30a-640f1d40d38b

We present radiative transfer models of the circumstellar environment
of classical T~Tauri stars, concentrating on the formation of the
H$\alpha$ emission. The wide variety of line profiles seen in
observations are indicative of both inflow and outflow, and we
therefore employ a circumstellar structure that includes both
magnetospheric accretion and a wind. We adopt two kinematic models for
the outflow, a generic bipolar wind and a disc wind. We perform
systematic investigations of the model parameters for the wind and the
magnetosphere to search for possible geometrical and physical
conditions which lead to the types of profiles seen in
observations. We find that both wind models can reproduce the wide
range profile types seen is observations, and that the most common
profile types observed occupy a large volume of parameter
space. Conversely, the most infrequently observed profile morphologies
require a very specific set of models parameters.  Although a
particular outflow model (bipolar or disc-wind) is not
well-constrained by a single line profile observations, we find that
the inclination dependence of the line equivalent width predicted by
the bipolar wind model agree with trends seen in the observation, but
the disc-wind model does not. Using the model results, we examine the
H$\alpha$ spectroscopic classification used by Reipurth et.~al, and
discuss the basic physical conditions that are required to reproduce
the profiles in each classified type.

@@

%Prior to the MA paradigm, many alternative models had
%been considered to explain the observed spectroscopic features mentioned
%above. For example, (1)~an Alfv\'{e}n wave-driven wind model
%(e.g.~\citealt{decampli:1981}; \citealt*{hartmann:1982}), (2)~a turbulent
%boundary layer (between the accretion disc and stellar surface) model
%(e.g.~\citealt{bertout:1988}; \citealt{basri:1989}), (3)~a chromospheric
%model (e.g.~\citealt*{calvet:1984}), and (4)~a disc wind model (e.g.
%\citealt*{calvet:1992b}; \citealt{kwan:1995}).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \begin{figure*}

%%   \begin{center}

%%     \includegraphics[%
%%       scale=0.5]{figures/overview3.eps}

%%   \end{center}

%%   \caption{The basic model configuration. The system consists of four
%%     components: (1)~the continuum source located at the origin $\left( o
%%     \right)$ of the cartesian coordinates $\left(x,y,z\right)$ -- the
%%     $y$-axis is into the paper, (2)~the magnetospheric accretion flow,
%%     (3)~the bipolar wind outflow, and (4)~the accretion disc. The density
%%     distribution is symmetric around the $z$-axis. The innermost radius of
%%     the magnetosphere (at the equatorial plane) coincides with the the
%%     inner radius of the accretion disc. From the innermost part of the
%%     accretion disc, the gas falls freely, moving along the magnetic field
%%     onto the surface of the star. The bipolar wind starts from just
%%     outside the largest radius 
%%     ($R_{\mathrm{mo}}$) of the magnetosphere (dotted line). }

%%   \label{fig:config_dipolar}

%% \end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}

    \includegraphics[%
      scale=0.3]{figures/overview3_alt.eps}

  \caption{The basic model configuration. The system consists of four
    components: (1)~the continuum source located at the origin $\left( o
    \right)$ of the cartesian coordinates $\left(x,y,z\right)$ -- the
    $y$-axis is into the paper, (2)~the magnetospheric accretion flow,
    (3)~the bipolar wind outflow, and (4)~the accretion disc. The density
    distribution is symmetric around the $z$-axis. The innermost radius of
    the magnetosphere (at the equatorial plane) coincides with the the
    inner radius of the accretion disc. From the innermost part of the
    accretion disc, the gas falls freely, moving along the magnetic field
    onto the surface of the star. The wind density is enhaced toward
    the polar directions to mimic the collimated wind structure of the
    magneto-hydrodynamical simulations (e.g.~\citealt{krasnopolsky:2003}). }

  \label{fig:config_dipolar}

@@

In order to understand how the different parts of the CTTS
circumstellar environment contribute to the formation of H$\alpha$,
the model space is divided into four different regions: (1)~a central
continuum source, (2)~the magnetospheric accretion flow, (3)~the wind
outflow, and (4)~the accretion disc. Fig.~\ref{fig:config_dipolar}
depicts the relative location of the regions in the model space.  The
density is assumed to be rotationally symmetric around the
$z$-axis. The innermost radius of the magnetosphere at the equatorial
plane coincides with the the inner radius of the accretion disc. From
the innermost part of the accretion disc, the gas falls freely, moving
along the magnetic field onto the surface of the star.  In the
following subsections we describe the details of model components.

@@

\subsection{Outflow geometries}

The magneto-centrifugal wind paradigm, first proposed by
\citet{blandford:1982}, has been often used to model the large-scale
wind structure of T Tauri stars, or the observed optical jets
(e.g.~HH~30 jet by \citealt{burrows:1996}; \citealt{ray:1996}). The
launching of the wind from a Keplerian disc is typically done by
treating the equatorial plane of the disc as a mass-injecting boundary
condition (e.g., \citealt{shu:1994}; \citealt{ustyugova:1995};
\citealt{ouyed:1997}; \citealt{krasnopolsky:2003}).  Depending on the
location of the open magnetic fields anchored to the disc, two
different types of winds are produced. If the field is constrained to
be near the co-rotation radius of stellar magnetosphere, an
{}``X-wind'' \citep{shu:1994} is produced. If the open field lines are
located in a wider area of the disc, a {}``disc-wind'' similar to that
of \citet{koenigl:2000} is produced \citep{krasnopolsky:2003}.  Recent
reviews on the jet/wind-disc connection can be found in
\citet{koenigl:2000} and \citet{pudritz:2005}.

Clearly there are several alternative outflow scenarios. Here we adopt
two simple kinematical models which broadly represent the results of
MHD simulations. The first is a bipolar wind model, in which the
outflow is collimated along the rotational axes (a model similar to
that of \citet{krasnopolsky:2003} in large scale, $>10$~au). The
second model is based on the disc-wind paradigm, in which the density
of the outflow increases towards the equatorial plane. The difference
in latitudinal density structure between the two outflow models should
provide an indication as to whether line profiles alone can be used to
constrain the outflow geometry. In the following sections we present
details of the two outflow structures.

\subsubsection{The bipolar wind}

\label{sub:wind_jet}

    \begin{tabular}{cc}
      \includegraphics[%
	clip,
	scale=0.45]{plots/discwind_vr_pol.eps}~~~~~~&
      \includegraphics[%
	clip,
	scale=0.45]{plots/density_wind_pol.eps}\tabularnewline
    \end{tabular}

  \end{center}
  \caption{The dependency of the bipolar wind velocity and density
    structure on the wind acceleration parameter $\beta$. The radial
    component of the wind velocity
    (equation~\ref{eq:wind_velocity_radial_part}) in polar direction,
    $\theta = 0$ as a function of radius is shown on the left. The wind
    density (equation~\ref{eq:wind_density_final} along polar direction as
    a function of radius is shown on the right. For all $\beta$ values, the
    density is normalised to the density at the base of the wind in polar
    direction, i.e. $\rho_0 = \rho\left(R_\mathrm{mo},0\right)$.  The
    smaller the value of $\beta$, the faster the acceleration of the
    wind. For the radii $r/R_{\mathrm{mo}} < 10$, both velocity and
    density are sensitive to the value of $\beta$. For all $\beta$ values,
    the radial velocity of the wind reaches the terminal velocity ($\sim
    210\,\kmps$) by $r/R_{\mathrm{mo}} \sim 1000$. The initial velocity
    $V_{0}=10\,\kmps$, which approximately corresponds to the thermal
    velocity of a hydrogen atom at 7500~K, is used for all $v_{r}$
    plots.}

  \label{fig:wind_vr_rho}

In a similar fashion to the simple model of \citet{appenzeller:2005},
the following parametrisation of collimated bipolar wind (region 3 in
Fig.~\ref{fig:config_dipolar}) is adopted. The wind velocity field,
$\boldsymbol{v}_{\mathrm{wind}}$, consists of radial and azimuthal
components which depend on the spherical coordinates $r$ and
$\theta$. The radial component $v_{r}\left(r\right)$ is assumed to
follow the classical beta-velocity law \citep[c.f.][]{castor:1979},
and the azimuthal component $v_{\phi}$ is assumed to be a constant
fraction ($\gamma$) of the Keplerian velocity for a given distance
($w=\sqrt{x^{2}+y^{2}}$) from the symmetry axis ($z$-axis) , i.e.:
\begin{equation}
     \boldsymbol{v}_{\mathrm{wind}}  = 
     v_{r}\hat{\boldsymbol{r}}+v_{\phi}\hat{\boldsymbol{\bphi}}
  \label{eq:wind_velocity_def}
\end{equation}
where 
\begin{equation}
   v_{r}\left(r\right) =
   v_{r0}+\left(v_{\infty}-v_{r0}\right)\left(1-\frac{R_{\mathrm{mo}}}{r}\right)^{\beta}\,,
   \label{eq:wind_velocity_radial_part}
\end{equation}
and
\begin{equation}
      v_{\phi}\left(w\right)=\gamma\,\left(\frac{GM_{*}}{w}\right)^{1/2}\,.
   \label{eq:wind_velocity_azimuth_part}
\end{equation}
Note that the base of the wind starts at $r=R_{\mathrm{mo}}$. This
naturally leads to a density discontinuity, but allows a simple
parameterisation of the wind whereas a self-consistent interface
between the magnetosphere and wind would require a full MHD
calculation that is beyond the scope of this paper. The range of polar
angle for the wind is restricted to
$\theta>\left|\theta_{\mathrm{disc}}\right|$ where
$\theta_{\mathrm{disc}}$ is the opening angle of the accretion
disc. $v_{r0}$ is the small radial velocity at the base of the wind
($r=R_{\mathrm{mo}}$). Normally, $v_{r0}=10\,\mathrm{km\, s^{-1}}$
which is approximately equal to the thermal velocity of hydrogen with
$T=7500\,\mathrm{K}$. Following \citet{appenzeller:2005},
$\gamma=0.05$ is adopted. The dependency of $v_{r}$ in polar direction
($\theta=0$) on the values of wind acceleration parameter $\beta$ is
shown in Figure~\ref{fig:wind_vr_rho}. All other parameters describing
the wind are fixed as the reference values given in
Table~\ref{tab:std_parameters}.

The density of the wind is assumed to be axi-symmetric and
separable in $r$ and $\theta$, 
\begin{equation}
    \rho\left(r,\theta\right)=P\left(r\right)\, F\left(\theta\right)
    \label{eq:wind_density_def}
\end{equation}
with 
\begin{equation}
    F\left(\theta\right)=n\,\cos^{b}\theta
    \label{eq:wind_density_theta}
\end{equation}
where $b$ is normally a positive even number (for density symmetric
about the equatorial plane), and $n$ is the angular normalisation
constant. For $b=0$, the wind is spherically symmetric except for
the parts disrupted by the accretion disc. The larger the value of
$b$, the higher the degree of the collimation. By integrating equation~\ref{eq:wind_density_theta}
over angles and normalising the integral to $4\pi$, one finds 
\begin{equation}
  n=\frac{1+b}{1-\cos^{1+b}\theta_{\mathrm{wind}}}\,.
  \label{eq:wind_density_theta_norm}
\end{equation}
Assuming the total mass-loss rate via the wind is
$\dot{M}_{\mathrm{wind}}$ and the mass-flux is constant, the radial
part of the density function is reduced to
$P\left(r\right)=\dot{M}_{\mathrm{wind}}\,\left[4\pi
r^{2}v_{r}\left(r\right)\right]^{-1}$;
hence, equation~\ref{eq:wind_density_def} becomes 
\begin{equation}
  \rho\left(r,\theta\right) = 
  \frac{n\,\dot{M}_{\mathrm{wind}}\,\cos^{b}\theta}{4\pi r^{2}\,
    v_{r}\left(r\right)}\,.
  \label{eq:wind_density_final}
\end{equation}
For a given mass-accretion rate, the wind mass-loss rate in our
typical model is assigned from the ratio of mass-loss to
mass-accretion rate
$\dot{M}_{\mathrm{wind}}/\dot{M}_{\mathrm{acc}}\approx0.1$, a value
indicated by both observations and MHD calculations
(e.g. \citealt{koenigl:2000}).  Figure~\ref{fig:wind_vr_rho} shows the
density along a streamline in the polar direction ($\theta=0$), for
different values of $\beta$ with all other parameters fixed at the
reference values (Table~\ref{tab:std_parameters}).  The density is
relatively sensitive to the value of $\beta$ for $r<10\,
R_{\mathrm{mo}}$.

\subsubsection{The disc wind}

\label{sub:Alternative-model-disc-wind}

@@

\subsection{Bipolar wind only models}

\label{sub:model-wind}

As mentioned above, the MA model alone
cannot explain the observations. As an
alternative model, we now examine formation of H$\alpha$ in a simple
bipolar wind as described in section \ref{sub:wind_jet}. The basic
wind model parameters introduced earlier are $\gamma$, $\beta$,
$v_{\infty}$, $b$, $\theta_{\mathrm{wind}}$, $T_{\mathrm{wind}}$ and
$\dot{M}_{\mathrm{wind}}$. We fix several of these parameters
at plausible values:  $\gamma=0.05$, $v_{\infty}=210\,\kmps$
\citep{appenzeller:2005} and
$\dot{M}_{\mathrm{wind}}=10^{-8}\,\mathrm{M_{\sun}\, yr^{-1}}$.
Further, the angular extent of the wind (the opening angle of the
wind) is assumed to be $\theta_{\mathrm{wind}}=80^{\circ}$. The degree
of collimation is initially chosen to be $b=4$, i.e. the ratio of the
density in polar direction to the density at the edge of the wind
closest to the accretion disc is about
$\left(\cos0^{\circ}/\cos80^{\circ}\right)^{4}\approx10^{3}$
Although the line is potentially
sensitive to the temperature structure of the wind, determination of a
self-consistent wind temperature is beyond the scope of this
paper. Readers are refereed to \citet{hartmann:1982} in which the wind
temperature structure is determined by balancing the radiative cooling
rate (assuming optically thin) with the MHD wave heating rate.
We pragmatically assume that the wind is isothermal at $T_{\mathrm{wind}}$.

Initially we examine the characteristics of the
H$\alpha$ profile  as a function of the wind acceleration parameter
$\beta$ and the isothermal wind temperature
$T_{\mathrm{wind}}$.  Figure~\ref{fig:atlas_wind_model} shows profiles computed at  $i=55^{\circ}$
 for a range of wind temperatures
between $7000\,\mathrm{K}$ and $9000\,\mathrm{K}$, and  $\beta$ between $0.2$ and $2.0$.

Classical P~Cygni profiles are prominent
in the models with lower values of $T_{\mathrm{wind}}$ and $\beta$
(a colder and faster accelerating wind). As the value of $\beta$
increases, the position of the blue-shifted absorption component moves
toward the line centre -- this can be seen clearly in the
$T_{\mathrm{wind}}=7000\,\mathrm{K}$ models. The velocity position
 of the absorption component of the fastest accelerating wind
($\beta=0.2$) is almost at the terminal velocity ($v_{\infty}\sim
210\,\kmps$), that for the slowest wind acceleration model
($\beta=4.0$) is located close the line centre ($\sim 50\,\kmps$).

 As
one can see in Figure~\ref{fig:wind_vr_rho}, the column density of the
gas moving at $v_{\infty}$ increases as the wind acceleration becomes
larger since the gas moving at $v_{\infty}$ becomes closer to the
photosphere. At the location where the absorption occurs, the gas is
moving faster for the models with the faster wind acceleration models;
hence, the location of the absorption component is bluer as $\beta$
becomes smaller (as the wind acceleration faster). The figure also
shows that the intensity of the emission component increases as
$\beta$ increases for a given $T_{\mathrm{wind}}$. This is mainly
because the density near the photosphere increases as the value of
$\beta$ increases (Figure~\ref{fig:wind_vr_rho}); hence, the
emissivity of increases. Unlike the magnetospheric accretion only
models (Figure~\ref{fig:atlas_acc_models}), the inverse P~Cygni
profiles are not seen here, nor the blue-ward asymmetry. Because of
the absorption component in P~Cygni profile, most of the profiles show
red-ward asymmetry except for the low temperature wind models with
high inclinations (not shown here).

    \includegraphics[%
      clip,
      scale=0.9]{figures/atlas_ha_i55_wind_only_final.eps}

  \caption{H$\alpha$ model profiles computed with wide ranges of the
    isothermal wind temperature ($T_{\mathrm{wind}}$) and the wind
    acceleration rate ($\beta$). The profiles are computed with the wind
    component only (no magnetospheric accretion flow or accretion
    disc). The `reference' model parameters (Table~\ref{tab:std_parameters}) are
    applied. The wind mass-loss rate ($\dot{M}_{\mathrm{wind}}$),
    the inclination ($i$) and the degree of columniation ($b$) are
    fixed at $10^{-8}~{\mathrm{M_{\sun}~yr^{-1}}}$, $55^{\circ}$ and
    $4.0$\, respectively. The temperature (indicated along the vertical
    axis) of the model increases from top to bottom, and the wind
    acceleration rate increases from left to right. The P~Cygni profile
    are seen for lower $T_{\mathrm{wind}}$ and lower $\beta$ (slow
    acceleration) models.  The position of the absorption the component
    moves toward the line centre as $\beta$ increases. The effect of
    line broadening becomes more prominent for higher $T_\mathrm{wind}$
    and higher $\beta$ models.}

  \label{fig:atlas_wind_model}

The dependency of the profile on the wind anisotropy parameter $b$
is demonstrated in Figure~\ref{fig:effect_wind_b_param}. Three models
presented in the figure have the same mass-loss rate ($10^{-8}\,\mathrm{M_{\sun}\, yr^{-1}}$),
and are computed at inclination angle $i=55^{\circ}$. The P~Cygni
profile feature becomes more prominent, and the line flux becomes
smaller as the wind becomes more isotropic (smaller $b$ value). For
a fixed mass-loss rate, the column density of along the line of a
sight to the observer at a medium inclination angle (e.g.~$i=55^{\circ}$)
increases as $b$ decreases (c.f.~equation~\ref{eq:wind_density_final});
hence, causing more absorption. Further more, the value of the highest
density in the wind becomes smaller for a smaller $b$ model. As the
wind becomes more isotropic, the material in the
polar direction has to be spread in the wider angular angles. This
results in a smaller highest density region in the wind; hence the
smaller line emissivity. For the same reason, the emission in the
wings is weaker for the model smaller $b$ value. We also note that
the asymmetry around the line centre decreases as the wind becomes
more isotropic.

The effect of varying the wind mass-loss rate is shown in Figure~\ref{fig:effect_wind_mdot}.
The line is very sensitive to the mass-loss rate. With all other parameters
fixed constant, $\dot{M}_{\mathrm{wind}}$ acts as an scaling factor
of the wind density (equation~\ref{eq:wind_density_final}). The
important processes in forming H$\alpha$ emission are the
recombination process and the subsequent downward transition to $n=3$
level. Assuming the ionisation fraction remains constant, the
recombination rate is proportional to the square 
of wind density. We have found the line flux of the models in the
figure is approximately proportional to the square of the mass-loss
rate. This very sensitive nature of H$\alpha$ would make the line
a good candidate for a mass-loss rate indicator; however, without
solving the ambiguity of the emission component from the magnetospheric
accretion flow and the wind, and without knowing the detailed temperature
structure of the wind, it could provide us a misleading a mass-loss
rate estimate of a T Tauri star.

    \includegraphics[%
      clip,
      scale=0.62]{figures/windonly_B_effect2.eps}

  \caption{The effect of wind collimation. The line profiles computed
    with with $\left(T_{\mathrm{wind}},~\beta \right) = \left(
    9000\Kelvin, 0.5\right)$ for the wind anisotropy parameter $b=0, 2, 4$
    are shown in dashed, dotted, and solid respectively.  All the other
    parameters are same as in Figure~\ref{fig:atlas_wind_model}.  The
    P~Cygni profile feature becomes prominent for the spherical wind model
    since the column density of along the line of a sight of the observer
    at a mid inclination angle ($i=55^{\circ}$) increases for lower $b$
    values for a fixed mass-loss rate of the wind. The blue-ward asymmetry
    is still present even in the isotropic wind case ($b=0$).}

  \label{fig:effect_wind_b_param}

    \includegraphics[%
      clip,
      scale=0.45]{figures/windonly_mdot_effect.eps}

  \caption{The effect of the wind mass-loss rate.  The profiles computed
    with $\left(T_{\mathrm{wind}},~\beta \right) = \left( 9000\Kelvin,
    0.5\right)$, and $\dot{M}_{\mathrm{wind}} = 1.0 \times
    10^{-8}~\mathrm{M_{\sun}\,yr^{-1}}$\, (solid), $0.5 \times
    10^{-8}~\mathrm{M_{\sun}\,yr^{-1}}$\, (dotted), and $0.25 \times
    10^{-8}~\mathrm{M_{\sun}\,yr^{-1}}$\, (dashed) are shown above. The
    models are computed with only the wind component (no magnetospheric
    accretion flow or accretion disc). With all other parameter fixed
    constant, the mass-loss rate behaves as a scaling factor for the wind
    density (equation~\ref{eq:wind_density_final}).}

  \label{fig:effect_wind_mdot}

\end{figure}

\subsection{Disc $+$bipolar wind}

\label{sub:model-disc-wind}

Next, we examine the effect of the size of the inner radius of the
disc and the distribution of the H$\alpha$ emission region. Figure~\ref{fig:effect_alpha_disc}
shows the wind emission profiles computed with and without the accretion
disc at $i=55^{\circ}$. The H$\alpha$ images (projected on to the
plane perpendicular to the observer at the location of the star) are
also shown in the same figure. Both models use the same parameters
as in Figure~\ref{fig:atlas_wind_model}, but with $\beta=0.5$ and
$T_{\mathrm{wind}}=9000\,\mathrm{K}$). The model without the disc
can be also considered as a model with a large inner radius of the
accretion disc, in which none of the wind emission region is blocked
by the disc --- provided the inclination angle is not high and
the disc is not or slightly flared.

The bipolar nature of the wind is clearly seen in the images. In this
example, the most of the wind emission occurs within a few to a several
(depending mainly on the value of $\beta$) stellar radii from the
star. The figure also shows that the accretion disc blocking the wind
emission from the bottom half of the wind which is moving away from
the observer. This results in the reduction of the wind emission in
the red wing. The image (mid panel) also shows that the accretion
disc is not occulting the stellar disc at this inclination; therefore,
the ratio of the wind emission flux to the photospherical continuum
flux becomes slightly smaller for the model with the disc. This results
in the reduction in the peak flux and the flux in the wings in the
normalised profile shape. If the disc becomes more flared (if the
opening angle exceeds $45^{\circ}$at the outer edge of the accretion
disc), the occultation of the stellar disc by the accretion could
be seen at this inclination angle. Although it is not shown here,
the continuum level of the two models are identical. For the models
with $i<30^{\circ}$, the presence of the disc has little/no effect
on the profile shapes. In the figures, the extent of the H$\alpha$
emission is smaller ($\sim 0.2\,\mathrm{au}$ or $\sim 20\, R_{*}$)
compared to the spectro-astrometric observations of \citet{takami:2003}
for RU~Lup and CS Cha which show 1--5~au scale outflows%
\footnote{\citet{takami:2003} their observation also show that some objects
(e.g~Z CMa and AS~353A) displaying the outflows of larger scale
($>50\,\mathrm{au}$); however, this could be formed by shocks rather
than MHD-wave heating (e.g. \citealt{hartmann:1982}) or X-ray heating
(e.g.~\citealt{shang:2002}). %
}; however, the extent of the line emission region becomes slightly
larger ($\sim0.4$~au) if a slower wind acceleration rate ( $\beta=4$)
is used.

Although not shown here, we have also computed the H$\alpha$ profiles
from the combination of the collimated wind and the accretion disc
($R_{\mathrm{di}}=2.2R_{*}$ and $R_{\mathrm{do}}=100\,\mathrm{au}$)
for the same range of $\beta$ and $T_{\mathrm{max}}$ values used in
Figure~\ref{fig:atlas_wind_model}.  As expected
(c.f.~Figure~\ref{fig:effect_alpha_disc}), the models show the
decrease in the flux in the blue wing when compared to the flux of the
same model computed without the accretion disc.  The overall behaviour
of the line profile shapes as functions of $\beta$ and
$T_{\mathrm{wind}}$ are very similar to that for the models without
the accretion disc.  A significant reduction of the flux in the red
wing is seen in the wind model with the accretion disc when compared
with the wind only model. The wings appear to be weaker in these
models compared to the wind only models because the ratio of the wind
emission flux to the continuum flux is lower.

    \begin{tabular}{ccc}
      \includegraphics[%
	clip,
	scale=0.3]{figures/disc_nodisc_compare.eps}&
      \includegraphics[%
	scale=0.3,
	angle=270,
	origin=lt]{figures/wind_thindisc.eps}&
      \includegraphics[%
	scale=0.3,
	angle=270,
	origin=lt]{figures/wind_nodisc.eps}\tabularnewline
    \end{tabular}

  \caption{A comparison of the wind models with and without the
    optically-thick and geometrically-thin disc.  The latter can be also
    equivalent to a model with a large inner radius of the accretion disc
    in which none of the wind emission region is not blocked provided the
    inclination is moderate. The left panel shows the profile computed
    with $i=55^{\circ}$, $\beta=0.5$ and $T_{\mathrm{wind}} =
    9000~\mathrm{K}$ (see text for other parameters used). The flux in the
    red wing of the line profile from the model with the disc (solid) is
    greatly reduced by the presence of the accretion disc compared to that
    from the model without the disc (dashed).  The peak flux and the flux
    in the wings are also reduced in the model with the disc because the
    disc does block the wind emission region, but none of the stellar
    disc, i.e. the continuum flux level is unchanged.  The H$\alpha$
    image (project in the plane of the observer) for the model with the
    disc (middle panel) and that for the model without (right plane) are
    also shown. The image flux is computed by integrating the specific
    flux over the wavelength range ($-800\,\kmps < v < 800\,\kmps$. The
    flux is in an arbitrary units and on a logarithmic (base $10$) scale. }

  \label{fig:effect_alpha_disc}

    \includegraphics[%
      clip,
      scale=0.45]{figures/summary.eps}

  \caption{A comparison of the profiles computed with the magnetosphere,
    the wind, the disc, and their combinations. The model configurations 
    used above are: (1)~magnetosphere only (dash), (2)~wind and disc
    (dot), and (3)~disc, wind and magnetosphere (solid).  The parameters
    used here are $T_{\mathrm{max}}=7500\,\Kelvin$,
    $\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\,yr^{-1}}$,
    $\dot{M}_{\mathrm{wind}}=10^{-8}\,\mathrm{M_{\sun}\,yr^{-1}}$,
    $T_{\mathrm{wind}}=9000\,\Kelvin$, and $\beta=0.5$. The profile for
    the case 3 is can be understood as a simple flux sum of the profiles
    from case 1 and 2.}

  \label{fig:summary_beta05_t8000}

@@

\subsection{Magnetosphere $+$ disc $+$bipolar wind}

\label{sub:model-acc-disc-wind}

We examine the characteristics of the line profiles arising from the
combination of the magnetospheric accretion flow, the wind, and the
accretion disc. For simplicity, the parameters for the magnetosphere
($T_{\mathrm{max}}=7500\,\mathrm{K}$ and
$\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$ are
fixed as in Figure~\ref{fig:atlas_acc_models}), and the wind mass-loss
rate, the isothermal wind temperature and the wind anisotropy
parameter to $\dot{M}_{\mathrm{wind}}=10^{-8}\,\mathrm{M_{\sun}\,
yr^{-1}}$
(i.e.~$\dot{M}_{\mathrm{wind}}=0.1\,\dot{M}_{\mathrm{acc}}$),
$T_{\mathrm{wind}}$ and $b=4$ respectively. The profile computed for
$i=55^{\circ}$ is given in Figure~\ref{fig:summary_beta05_t8000},
along with the profile computed with the magnetosphere only and that
with the wind plus accretion disc configurations for a comparison. As
one can see from the figure, the profile computed for the
disc--wind--magnetosphere hybrid model can be understood as a simple
flux sum of the magnetospheric component and the wind component. In
this example, the wind causes the absorption of the blue side of the
profile, and it adds extra flux around the line centre. As seen
before, the presence of the disc causes a reduction in the red side of
the profile since the emission from the receding gas is blocked by the
disc. The resulting profile appears to be narrower than that of the
magnetosphere model.

    \includegraphics[%
      clip,
      scale=0.9]{figures/atlas_ha_i55_wind_acc_alpha_final2.eps}

  \caption{Same as Figure~\ref{fig:atlas_wind_model}, but the models
    include the magnetospheric accretion flow (with
    $T_{\mathrm{\max}}=7500~\mathrm{K}$ and
    $\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\,yr^{-1}}$),  and the
    accretion disc. The mass-loss rate of the wind is
    $\dot{M}_{\mathrm{wind}}=10^{-9}\,\mathrm{M_{\sun}\,yr^{-1}}$. A
    relatively small mass-loss rate is chosen to avoid unrealistically
    strong P-Cyg absorption component (see text in
    section~\ref{sub:model-acc-disc-wind} and
    Figure~\ref{fig:effect_mdot_disc_wind_acc}).  For a comparison, the
    profile computed with only the magnetospheric accretion flow (dotted)
    is overlayed. All the profiles are computed at the inclination angle
    $i=55^{\circ}$. As $\beta$ becomes larger, the position of the wind
    absorption component moves toward the line centre. Although little
    wind emission is seen in this temperature range, it becomes more
    important in a model with a higher temperature and a larger wind
    mass-loss rate. The profiles resembles the observed H$\alpha$ emission
    profiles affected by the wind.}

  \label{fig:atlas_disc_wind_acc_model}

The effect of changing the ratio of the wind mass-loss rate to
mass-accretion rate
($\mu=\dot{M}_{\mathrm{wind}}/\dot{M}_{\mathrm{acc}}$) is demonstrated
in Figure~\ref{fig:effect_mdot_disc_wind_acc}. With a fixed value of
mass-accretion rate
($\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$), the
mass-loss rate is varied. The figure shows that as $\mu$ increases the
P-Cygni absorption deepens, and the position of minimum flux in the
absorption trough appears to move blue-ward.  On the other hand, the
emission component becomes stronger as $\mu$ increases. With this
combination of the parameters for the magnetosphere and the wind, the
absorption component tends to be too strong for the models $\mu>0.025$
(with this model configuration), at all inclinations (not shown here),
compared with the observed H$\alpha$ profiles
(e.g.\citealt{reipurth:1996}; \citealt{alencar:2000}).
%As we will see later (section~\ref{fig:ttau_model}),
%$\mu$ can be as large as $\sim0.3$ with different parameters of
%the magnetosphere.

Again, the H$\alpha$ profiles with the same range of the wind
temperature ($T_{\mathrm{wind}}$) and the wind acceleration parameter
($\beta$) used in Figure~\ref{fig:atlas_wind_model} are computed, but
this time with $\mu=0.01$.  The parameters of the magnetosphere are
fixed as in Figure~\ref{fig:acc_inclination_effect}. The results are
shown in Figure~\ref{fig:atlas_disc_wind_acc_model} along with the
model with only the magnetosphere for a comparison. A relatively small
value of $\mu$ is chosen to avoid absorption component of the P-Cyg
profile becoming too strong, as seen in
Figure~\ref{fig:effect_mdot_disc_wind_acc}.  In general, even with
small values of the anisotropy parameter ($b=0$ and $2$), the
absorption component is found to be too strong when $\mu=0.1$ is used
with this combination of the magnetosphere and the wind. The figure
shows that the position of the wind absorption component moves toward
the line centre as $\beta$ becomes larger.  Although little wind
emission is seen in this temperature range, it becomes more important
in a model with a higher temperature and a larger wind mass-loss
rate. In fact, the model with $\beta=2.0$ and
$T_{\mathrm{wind}}=9000\,\mathrm{K}$ displays a small amount of the
wind emission at $V\sim50\,\mathrm{km\, s^{-1}}$.

As seen in Figure~\ref{fig:summary_beta05_t8000}, the profiles shown
here can be understood as the combination of the emission from the
magnetosphere and the absorption of the flux by the wind component.
In general, the emission from the accretion disc is negligible at
H$\alpha$ wavelength, but the absorption of the continuum flux from
the photosphere by the disc is important for predicting correct line
strengths, especially at a high inclination.

>>>>

We present radiative transfer models of the circumstellar environment
of classical T~Tauri stars, concentrating on the formation of the
H$\alpha$ emission. The wide variety of line profiles seen in
observations are indicative of both inflow and outflow, and we
therefore employ a circumstellar structure that includes both
magnetospheric accretion and a disc wind.  We perform
systematic investigations of the model parameters for the wind and the
magnetosphere to search for possible geometrical and physical
conditions which lead to the types of profiles seen in
observations. We find that the hybrid models can reproduce the wide
range profile types seen is observations, and that the most common
profile types observed occupy a large volume of parameter
space. Conversely, the most infrequently observed profile morphologies
require a very specific set of models parameters.  We investigate the
wind contribution to the line profile as a function of model parameters, and
examine the reliability of H$\alpha$ as a mass accretion diagnostic.
Finally, we examine the
H$\alpha$ spectroscopic classification used by Reipurth et.~al, and
discuss the basic physical conditions that are required to reproduce
the profiles in each classified type.

@@

@@

%%%%%%%%%%%%%%

@@

In order to understand how the different parts of the CTTS
circumstellar environment contribute to the formation of H$\alpha$,
the model space is divided into four different regions: (1)~a central
continuum source, (2)~the magnetospheric accretion flow, (3)~the wind
outflow, and (4)~the accretion disc. Fig.~\ref{fig:config_discwind}
depicts the relative location of the regions in the model space.  The
density is assumed to be rotationally symmetric around the
$z$-axis. The innermost radius of the magnetosphere at the equatorial
plane coincides with the the inner radius of the accretion disc. From
the innermost part of the accretion disc, the gas falls freely, moving
along the magnetic field onto the surface of the star.  In the
following subsections we describe the details of model components.

@@

\subsection{The disc-wind outflow geometru}
\label{sub:wind_jet}

The magneto-centrifugal wind paradigm, first proposed by
\citet{blandford:1982}, has been often used to model the large-scale
wind structure of T Tauri stars, or the observed optical jets
(e.g.~HH~30 jet by \citealt{burrows:1996}; \citealt{ray:1996}). The
launching of the wind from a Keplerian disc is typically done by
treating the equatorial plane of the disc as a mass-injecting boundary
condition (e.g., \citealt{shu:1994}; \citealt{ustyugova:1995};
\citealt{ouyed:1997}; \citealt{krasnopolsky:2003}).  Depending on the
location of the open magnetic fields anchored to the disc, two
different types of winds are produced. If the field is constrained to
be near the co-rotation radius of stellar magnetosphere, an
{}``X-wind'' \citep{shu:1994} is produced. If the open field lines are
located in a wider area of the disc, a {}``disc-wind'' similar to that
of \citet{koenigl:2000} is produced \citep{krasnopolsky:2003}.  Recent
reviews on the jet/wind-disc connection can be found in
\citet{koenigl:2000} and \citet{pudritz:2005}. Clearly there are several alternative outflow scenarios. Here we adopt
a simple kinematical model, based on the disc-wind paradigm, which broadly represents the results of
MHD simulations,

@@

@@
